They are Very Smart: Genetic-Algorithm-Based Pathfinding in Games

### 2018-12-04 paper.tex

At this time, we are currently waiting on actual results from our project, but we hope that given a large enough population of zombies, we might see populations of zombies evolving to counteract players' strategies, although we may also see players coming up with ways to counteract the zombies by exploiting elements of the fitness function and construction system in ways unforeseen, possibly by playing in such a way as to allow the zombies to optimize for something other than killing the player's home base. Another possibility we might expect to see is the development of swarm or anti-swarm behavior based on the effectiveness of player defenses, as well as zombies possibly hitting minima or maxima depending on hard-coded numbers in the game environment.

In the future, we would experiment with elements such as variable damage output from towers and other defenses (so as to avoid the aforementioned numeric minima and maxima), as well as coming up with ways to encourage attacking the player's base more as time goes on.
\begin{figure}
	\includegraphics[width=\columnwidth]{InGame2}
	\caption{An in-game image}
\end{figure}

>>>>

At this time, we are currently waiting on actual results from our project, but
we hope that given a large enough population of zombies, we might see
populations of zombies evolving to counteract players' strategies. Evolution
will likely have little trouble tackling a trivial or consistent player
strategy, like consistently rebuilding the same towers in the same place and
relying heavily on a single tower type, but evolution may have trouble keeping
up with a player that consistently changes strategy over the course of the game.
Our system of softmaxing the genome speed, health, and damage could also
result in progressively larger values relative to the mutation rate, making it
hard for long-evolved values to be changed in the face of changing player
strategies.

One approach to solving this problem could be moving towards more complicated
models of selection and reproduction than our current keep \(n\) percent,
asexually reproduce the best genomes and then do a straight pass of mutation.
Moving to a system where reproductive resources are distributed according to the
Z-score (standard deviations from the mean) or similar of fitness that give more
reproductive power to individual that more drastically improve on the status quo
could allow for a system that reacts more strongly to change. Another option
could be coding part of the mutation rate in genomes to allow evolution to
determine it's own speed. The two previous approaches could even be combined for
as dynamic an evolution environment as possible.

Sexual (genderless) reproduction could also be introduced as a way of trying to
increase diversity. Plain sexual reproduction without any further restrictions
could allow good traits that evolve in parallel to be combined into a single
individual with the benefits of both, which has fairly obvious advantages.
This simple approach to sexual reproduction could, however, force a single
species to dominate quickly depending on how well neural net parameters combine,
as it could easily be the case that two disparately evolving neural nets combine
into a total dud, despite having fit parents. One way around that would be to
condition reproduction on similarity, such genomes are more likely to reproduce
with those that are structurally similar. Another approach would be to allow
genomes to determine mating preference based on other genomes' content and
fitness, and then run them through a tournament to determine mates. This would
have the additional benefit of allowing additional emergent behavior that has
not been conceived of in this speculation.

Conceptually, evolving neural nets has the massive benefit of not worrying about
how to differentiate and nicely back-propagate

individuals that jump out ahead of the population could

Discussion: talk about the lack of direction vectors being fed into the neural nets.

may also see players coming up with ways to counteract the zombies by exploiting
elements of the fitness function and construction system in ways unforeseen,
possibly by playing in such a way as to allow the zombies to optimize for
something other than killing the player's home base. Another possibility we
might expect to see is the development of swarm or anti-swarm behavior based on
the effectiveness of player defenses, as well as zombies possibly hitting minima
or maxima depending on hard-coded numbers in the game environment.

In the future, we could experiment with elements such as variable damage output
from towers and other defenses (so as to avoid the aforementioned numeric minima
and maxima), as well as coming up with ways to encourage attacking the player's
base more as time goes on; this could include ways
